---
title: "MultiONet Analysis Workflow"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r include=FALSE}
mypath=""
deg<-read.csv("/GSE254185_kidney/data/DEG_per_cluster.csv")
head(deg)
table(deg$Cluster)

peak<-read.csv("/GSE254185_kidney/data/top_peak_per_cluster.csv")
head(peak)
table(peak$Cluster)
```

# Introduction

This vignette demonstrates the workflow using the **MultiONet** R package, specifically designed for network-based multi-omics integration and enrichment analyses. We will apply it to kidney data from Maximilian Reck et al. (Nature Communications, 2025) to explore the functionality of MultiONet.

## Required Packages

Install and load necessary libraries:

```{r}
library(dplyr)
library(magrittr)
library(MultiONet)
library(OmnipathR)
library(igraph)
library(GGally)
```


# Workflow

Define target clusters and initialize variables:

```{r}
target_clusters <- c("PT Injured", "PT Inflammatory")
cluster_a <- target_clusters[1]
```

# Data Preparation

Subset your DEG and peak data based on selected clusters:

```{r}
data_a <- deg %>% filter(Cluster == cluster_a, adj..p.value < 0.05)
data_b <- peak %>% filter(Cluster == cluster_a, adj..p.value < 0.05)
```

# Gene Set Enrichment Analysis (ORA)

Run overrepresentation analysis (ORA) on DEGs, peak genes, their intersection, and their union.

```{r}
Largest_of_deg_ora <- run_ora(data_a$Gene)
Largest_of_peak_ora <- run_ora(data_b$Closest.Gene)
Largest_of_intersection_ora <- run_ora(intersect(data_a$Gene, data_b$Closest.Gene))
Largest_of_combined_ora <- run_ora(union(data_a$Gene, data_b$Closest.Gene))
```

Examine the dimensions of the results:

```{r}
dim(Largest_of_deg_ora)
dim(Largest_of_peak_ora)
dim(Largest_of_intersection_ora)
dim(Largest_of_combined_ora)
```

# Compare Gene Lists

Compare DEG and peak gene lists to identify overlaps and differences. There is 8% overlap between the two gene sets.

```{r}
input <- list(deg = data_a$Gene, peak = data_b$Closest.Gene)
result <- compare_gene_list(input)
```

# Network Generation

Create networks based on the gene lists. Networks can be constructed from (i) differentially expressed genes (DEGs), (ii) genes associated with peaks, (iii) their intersection, and (iv) their union. For each case, the largest connected subnetwork can also be extracted.

```{r}
setwd(paste0(mypath,"/data"))

result <- make_networks(
  input,
  indirect_neighbors = 1,
  score_threshold = 996,
  network_file_prefix = cluster_a,
  string_link_file = "9606.protein.links.full.v11.5.txt.gz",
  string_alias_file = "9606.protein.aliases.v11.5.txt.gz",
  plot_network = TRUE,
  label_genes = TRUE,
  plot_network_node_threshold = 2000,
  label_size = 2
)
```

# ORA of Largest Networks

Perform ORA for genes in the largest networks then plot a upset plot to examine the similarity between the two results.:

```{r}
Largest_network_of_deg_ora <- run_ora(V(result[['largest_networks']][["Largest_network_of_deg_1"]])$name)
Largest_network_of_peak_ora <- run_ora(V(result[['largest_networks']][["Largest_network_of_peak_1"]])$name)

compare_gene_sets(list(set1 = Largest_network_of_deg_ora, set2 = Largest_network_of_peak_ora))
```

# Reducing Enrichment Results

Reduce redundancy in enrichment results using semantic similarity. The returned matrix has parent node information that can be used to remove duplicate results:

```{r}
Largest_network_of_deg_ora %<>% mutate(source = "deg")
Largest_network_of_peak_ora %<>% mutate(source = "peak")

combined_raw <- rbind(Largest_network_of_deg_ora, Largest_network_of_peak_ora)
Largest_network_of_deg_peak_ora_reduced <- reduce_enrichment_result(combined_raw)
Largest_network_of_deg_ora_reduced <- reduce_enrichment_result(Largest_network_of_deg_ora)
Largest_network_of_peak_ora_reduced <- reduce_enrichment_result(Largest_network_of_peak_ora)

head(knitr::kable(Largest_network_of_peak_ora_reduced))
```

# Heatmap Visualization

Create a heatmap to visualize reduced enrichment results. Heatmap is based on top 20 gene sets from each gene set results (DEG & peak). The common parents between the two gene sets and the unique parents of each gene set are kept in the result:

```{r}
reduced_heatmap_r <- reduced_heatmap(
  Largest_network_of_deg_peak_ora_reduced[['processed_result']],
  combined_raw,
  Largest_network_of_deg_ora_reduced[['processed_result']],
  Largest_network_of_peak_ora_reduced[['processed_result']],
  file_name="GSE254185_kidney",
  max_per_set = 20
)
```
```{r, out.width="100%"}
knitr::include_graphics(paste0(mypath,"/scripts/GSE254185_kidney.pdf"))
```


# Directed Network Visualization

Visualize directed networks with Omnipath, a database built from above 100 resources:

```{r}
make_directed_graph(result[['largest_networks']][["Largest_network_of_deg_1"]])
make_directed_graph(result[['largest_networks']][["Largest_network_of_peak_1"]])
```

# Centrality Analysis (PageRank)

Calculate and visualize PageRank to identify influential nodes:

```{r}
pagerank_result <- get_pagerank(result[['largest_networks']][["Largest_network_of_combined_1"]])
```

# Conclusion

This vignette illustrates the use of **MultiONet** to systematically analyze and visualize multi-omics data, enabling comprehensive biological insights through integrated network and enrichment analyses.
